---
title: "re_airquality"
author: "Flueckiger_Rubens; 21-118-385"
date: "2025-03-10"
output: 
  html_document:
    toc: true
---

# Report Excersice re_airquality

## 1 Load packages

```{r, message=FALSE, error=FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)
library(readr)
library(lubridate)
library(dplyr)
```

## 2 Dataset Information

The built-in dataset *airquality* has daily air quality measurements from New York between 1.5.1973 and 30.8.1973. The dataset contains:

-   Ozone [ppb]
-   Solar radiation [J/cm^2^] at frequency band between 4'000 and 7'700 Angstroms
-   Average windspeed [mph]
-   Maximum daily temperature [°F]

With the provided data, the effect of ozone production at near surface level is analysed. Ozone at near surface levels is created followingly:

$$NO_2 + O_2 <==> NO + O_3$$

Both reactions take place simultaneously. To enrich the concentration, emitted volatile organic compounds (VOC) can bond with NO and thus limiting the backward reaction.

I hypothesize that a higher wind speed, which should enhance mixing, and solar radiation leads to a higher ozone concentration.

## 3 Analysis

### 3.1 Monthly fluxes

First monthly fluxes are analysed by aggregating daily values over the month.

```{r, echo=FALSE, message=FALSE}
#Deleting each row with NA values and storing it into as separate variable 
oz <- airquality %>% drop_na()

#Aggregating over months
#Mean values
oz_mean <- oz %>% group_by(Month) %>% summarise(
  Ozone = round(mean(Ozone), 1),
  Wind = round(mean(Wind), 1),
  Solar.R = round(mean(Solar.R), 1),
  Temp = round(mean(Temp), 1)
)

#Minimum values
oz_min <- oz %>% group_by(Month) %>% summarise(
  Ozone = min(Ozone),
  Wind = min(Wind),
  Solar.R = min(Solar.R),
  Temp = min(Temp)
)

#Maximum values
oz_max <- oz %>% group_by(Month) %>% summarise(
  Ozone = max(Ozone),
  Wind = max(Wind),
  Solar.R = max(Solar.R),
  Temp = max(Temp)
)

#Variance
oz_var <- oz %>% group_by(Month) %>% summarise(
  Ozone = round(var(Ozone), 1),
  Wind = round(var(Wind), 1),
  Solar.R = round(var(Solar.R), 1),
  Temp = round(var(Temp), 1)
)

#Creating tables containing statistics of
#Windspeed
wind_stats <- data.frame(oz_min$Month, oz_min$Wind, oz_max$Wind, oz_mean$Wind, oz_var$Wind)
colnames(wind_stats) <- c("Month", "Minimum", "Maximum", "Mean", "Variance")

#Ozone
ozone_stats <- data.frame(oz_min$Month, oz_min$Ozone, oz_max$Ozone, oz_mean$Ozone, oz_var$Ozone)
colnames(ozone_stats) <- c("Month", "Minimum", "Maximum", "Mean", "Variance")

#Solar radiation
solar_stats <- data.frame(oz_min$Month, oz_min$Solar.R, oz_max$Solar.R, oz_mean$Solar.R, oz_var$Solar.R)
colnames(solar_stats) <- c("Month", "Minimum", "Maximum", "Mean", "Variance")

#Temperature
temp_stats <- data.frame(oz_min$Month, oz_min$Temp, oz_max$Temp, oz_mean$Temp, oz_var$Temp)
colnames(temp_stats) <- c("Month", "Minimum", "Maximum", "Mean", "Variance")

#Create table
knitr::kable(wind_stats, caption = "Windspeed statistics")
knitr::kable(ozone_stats, caption = "Ozone statistics")
knitr::kable(temp_stats, caption = "Temperature statistics")
knitr::kable(solar_stats, caption = "Solar radiation statistics")
```

Now four plots containing the range and mean value of the observed variables are created.

```{r, echo=FALSE, results="hide"}
#Expression solar radiation
sol_rad <- substitute("Solar radiation [J/"*cm^2*"]") 
    as.character(as.expression(sol_rad))
```

```{r, echo=FALSE, message=FALSE, fig.cap="Figure 1: Monthly evolutions", fig.align='center'}
#Monthly mean, max and min ozone values
p1 <- ggplot(data = oz_mean,aes(x = Month,y = Ozone)) + geom_line() + #Average line
  geom_line(data = oz_min, aes(x = Month, y = Ozone),col = "navy") + #Minimum line
  geom_line(data = oz_max, aes(x = Month, y = Ozone),col = "navy") + #Maximum line
  geom_ribbon(aes(x = Month, ymin = oz_min$Ozone, ymax = oz_max$Ozone),fill = "lightblue",alpha = 0.3) + #Filling 
  labs(title = "Monthly ozone variation & range") + #Title
  labs(y = "Ozone [ppb]") + #y-axis label
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Monthly mean, max and min temperature values
p2 <- ggplot(data = oz_mean,aes(x = Month,y = Temp)) + geom_line() + #Average line
  geom_line(data = oz_min, aes(x = Month, y = Temp),col = "red") + #Minimum line
  geom_line(data = oz_max, aes(x = Month, y = Temp),col = "red") + #Maximum line
  geom_ribbon(aes(x = Month, ymin = oz_min$Temp, ymax = oz_max$Temp),fill = "red",alpha = 0.2) + #Filling 
  labs(title = "Monthly Temperautre variation & range") + #Title
  labs(y = "Temperature [°F]") + #y-axis label
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Monthly mean, max and min wind values
p3 <- ggplot(data = oz_mean,aes(x = Month,y = Wind)) + geom_line() + #Average line
  geom_line(data = oz_min, aes(x = Month, y = Wind),col = "darkgreen") + #Minimum line
  geom_line(data = oz_max, aes(x = Month, y = Wind),col = "darkgreen") + #Maximum line
  geom_ribbon(aes(x = Month, ymin = oz_min$Wind, ymax = oz_max$Wind),fill = "lightgreen",alpha = 0.3) + #Filling 
  labs(title = "Monthly Wind speed variation & range") + #Title
  labs(y = "Wind speed [mph]") + #y-axis label
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Monthly mean, max and min solar radiation values
p4 <- ggplot(data = oz_mean,aes(x = Month,y = Solar.R)) + geom_line() + #Average line
  geom_line(data = oz_min, aes(x = Month, y = Solar.R),col = "darkorange") + #Minimum line
  geom_line(data = oz_max, aes(x = Month, y = Solar.R),col = "darkorange") + #Maximum line
  geom_ribbon(aes(x = Month, ymin = oz_min$Solar.R, ymax = oz_max$Solar.R),fill = "gold",alpha = 0.3) + #Filling 
  labs(title = "Monthly solar radiation variation & range") + #Title
  labs(y = sol_rad) + #y-axis label
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  theme(plot.title = element_text(hjust = 0.5)) #Center title

cowplot::plot_grid(p1,p2,p3,p4)
```

Already one can see for the months July and August, temperature, solar radiation and ozone concentration seem to be the highest.

In a next step we take a closer look at inner monthly variations with the help of violin plots.

```{r, echo=FALSE, fig.cap="Figure 2: Monthly distributions", fig.align='center'}
#Violin plot ozone
v1 <- ggplot(oz, aes(factor(Month), Ozone, fill = Month)) + 
  geom_violin() + #Generate violin plot
  scale_fill_viridis_c() + #Color palette
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  labs(x = "Ozone [ppb]") + #x-axis label
  labs(title = "Monthly ozone distribution") + #Title
  labs(y = "Month") + #y-axis label
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Violin plot temperature
v2 <- ggplot(oz, aes(factor(Month), Temp, fill = Month)) + 
  geom_violin() + #Generate violin plot
  scale_fill_viridis_c() + #Color palette
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  labs(x = "Temperature [°F]") + #x-axis label
  labs(title = "Monthly temperature distribution") + #Title
  labs(y = "Month") + #y-axis label
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Violin plot wind speed
v3 <- ggplot(oz, aes(factor(Month), Wind, fill = Month)) + 
  geom_violin() + #Generate violin plot
  scale_fill_viridis_c() + #Color palette
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  labs(x = "Wind speed [mph]") + #x-axis label
  labs(title = "Monthly wind speed distribution") + #Title
  labs(y = "Month") + #y-axis label
  theme(plot.title = element_text(hjust = 0.5)) #Center title

#Violin plot solar radiation
v4 <- ggplot(oz, aes(factor(Month), Solar.R, fill = Month)) + 
  geom_violin() + #Generate violin plot
  scale_fill_viridis_c() + #Color palette
  theme_bw() +
  theme(legend.position = "none") + #Hide legend
  labs(x = sol_rad) + #x-axis label
  labs(title = "Monthly solar radiation distribution") + #Title
  labs(y = "Month") + #y-axis label
  theme(plot.title = element_text(hjust = 0.5)) #Center title

cowplot::plot_grid(v1,v2,v3,v4)
```

One can see that the distributions for low wind speed and ozone, as well as solar radiation and temperature are very similar, which can be explained as solar radiation is the source of temperature.

### 3.2 Scatterplot

**Ozone vs. solar radiation**

Since wind speed seems to have a strong effect for low velocities, we first plot solar radiation against ozone concentration. Then a trend line including the regression coefficient is calculated.

```{r, echo=FALSE, results="hide"}
#linear ozone-solar radiation model
lm_os <- lm(Ozone ~ Solar.R,oz)               #Creating linear model

#Expression Linear function & R^2 value
eq <- substitute(y == a + b*x*";"~~R^2~"="~r2,
                 list(a = format(unname(coef(lm_os)[1]), digits = 2),
                      b = format(unname(coef(lm_os)[2]), digits = 2),
                      r2 = format(summary(lm_os)$r.squared, digits = 3)))
as.character(as.expression(eq))
```

```{r, echo=FALSE, message=FALSE, fig.cap="Figure 3: Ozone as a function of solar radiation", fig.align='center'}
#Plot regression
ggplot(data = oz,aes(x = Solar.R, y = Ozone)) + geom_point(shape = 20, size = 3) +
  geom_smooth(method = "lm",col = "red",se = F) + #Regression line
  theme_bw() +
  labs(y = "Ozone [ppb]") + #x-axis label
  labs(subtitle = eq) + #Subtitle with R^2
  labs(x = sol_rad) + #y-axis label
  labs(title = "Ozone as a function of solar radiation") + #Title
  theme(plot.title = element_text(hjust = 0.5)) #Center title
```

One can see that the regression coefficient has a low value of 0.113 which means that solar radiation only explains about 10% of the ozone concentration effect. 

**Ozone vs. solar radiation and wind speed**

By plotting the wind speed in colours, we can account for this effect as well.

```{r, echo=FALSE, message=FALSE, fig.cap="Figure 4: Ozone as a function of solar radiation and windspeed", fig.align='center'}
#Plot regression & wind
ggplot(data = oz,aes(x = Solar.R, y = Ozone, colour = Wind)) +
  geom_point(shape = 20, size = 3) +
  geom_smooth(method = "lm",col = "red",se = F) + #Regression line
  scale_color_viridis_c() +
  theme_bw() +
  labs(y = "Ozone [ppb]") + #x-axis label
  labs(x = sol_rad) + #y-axis label
  labs(title = "Solar radiation & Ozone & Wind") + #Title
  labs(subtitle = eq) + #Subtitle with R^2
  labs(colour = "Windspeed\n[mph]") + #Note: \n creates a new row
  theme(plot.title = element_text(hjust = 0.5)) #Center title
```

According to this new plot, we can now see that especially low wind speeds account for especially high ozone concentration.

## 4 Conclusion

According to the analysis done above, my hypothesis is correct regarding higher solar radiation leading to higher ozone concentrations. In contrast to my hypothesis, lower wind speed seams to produce more ozone, which may root from reaction velocity. For higher wind speeds, the reaction of VOC’s with NO may be too low to limit the NO concentration and therefore limiting the enrichment of ozone.
